### 2024-03-08 models/negative_bin_model.Rmd
Summary information
created tables and graphs showing the leading causes of death for overall population and by race

---
title: "negative_bin_model.Rmd"
format: html
---

```{r}
# load libraries
library(MASS)
library(dplyr)

# load files 
file_path <- "../scripts/death_count_by_Cause.csv"
death_count_by_cause <- read.csv(file_path)

death_count_by_cause$Leading.Cause <- as.factor(death_count_by_cause$Leading.Cause)

# Attempt to increase the iteration limit to allow convergence
nb_model <- glm.nb(Total_Deaths ~ Leading.Cause, 
                   data = death_count_by_cause, 
                   control = glm.control(maxit = 1000)) # Increase max iterations to 1000

# Check the summary
summary(nb_model)

```